---
title: "Ofsted Outcome Predictor Markdown"
author: "Andy Cartledge"
date: "18/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

This document will provide the testing ground for my Ofsted Outcome Predictor project.

First though we need to install some packages

```{r Install packages}

# This lists all the packages used in the project, more packages may be added as the project rolls on!

if(!require("tidyverse")) install.packages("tidyverse", dependencies = TRUE)
if(!require("openxlsx")) install.packages("openxlsx", dependencies = TRUE)
if(!require("stringr")) install.packages("stringr", dependencies = TRUE)
if(!require("pivottabler")) install.packages("pivottabler", dependencies = TRUE)
if(!require("devtools")) install.packages("devtools", dependencies = TRUE)
if(!require("rpart")) install.packages("rpart", dependencies = TRUE)
if(!require("rpart.plot")) install.packages("rpart.plot", dependencies = TRUE)
if(!require("tidymodels")) install.packages("tidymodels", dependencies = TRUE)
if(!require("broom.mixed")) install.packages("broom.mixed", dependencies = TRUE)
if(!require("readr")) install.packages("readr", dependencies = TRUE)
if(!require("rstanarm")) install.packages("rstanarm", dependencies = TRUE)
if(!require("ranger")) install.packages("ranger", dependencies = TRUE)

```

Then we need to load the packages into memory

```{r Load packages}

library(tidyverse)
library(openxlsx)
library(stringr)
library(pivottabler)
library(devtools)
library(rpart)
library(rpart.plot)
library(tidymodels)
library(broom.mixed)
library(readr)
library(rstanarm)
library(ranger)

```

Next I will load the excel file and the csv copy of edubase

```{r Load excel files}

# The data I need is on the 10th sheet and I'm also interested in dates. Also the top row is not the header row for the sheet, hence the different start row.

ofsted <- read.xlsx(
  "data/Management_information_-_state-funded_schools_-_as_at_31_August_2020.xlsx", 
  sheet = 10, 
  startRow = 2, 
  detectDates = TRUE, 
  na.strings = "NULL"
  )

edubase <- read_csv(
  "data/edubasealldata20200921.csv", 
  col_types = cols(`SEN13 (name)` = col_skip(), 
                   FEHEIdentifier = col_skip(), 
                   InspectorateReport = col_skip(), 
                   DateOfLastInspectionVisit = col_skip(), 
                   SiteName = col_skip()
                   ),
  guess_max = 10000
  )

```

Now I create a new table from the edubase dataset. This will contain all the fields I desire from that dataset, it will also create a new

```{r Code some academy's}

edubase_to_ofsted <- select(
  mutate(
    edubase, 
    Academy = ifelse(`EstablishmentTypeGroup (code)` == 10 | `EstablishmentTypeGroup (code)` == 11,1,0)),
    URN,
    Easting,
    Northing,
    Academy,
    `DistrictAdministrative (code)`,
    `DistrictAdministrative (name)`,
    `LSOA (name)`,
    `LSOA (code)`
)

```

```{r Match edubase data to Ofsted to add Academy column}

#This could likely be done within the ofsted data itself and I will attempt this after matching the data together

ofsted <- left_join(x = ofsted, y = edubase_to_ofsted, x.URN = y.URN)

```

```{r Add in LA code}

ofsted <- mutate(ofsted, LA.Code = as.integer(substr(LAESTAB,1,3)))

```

```{r Add in difference between inspection dates difference}

ofsted <- ofsted %>%
  mutate(days_between = Publication.date - Previous.publication.date)

```

Now that we have added the columns need from the edubase dataset we can now move onto an important step, removing schools which do not have a previous inspection. This will, most likely, be due to them being newly opened schools. Note that Ofsted recently changed its methodology to map new academys records to their predecessor schools records. However Free Schools and the like that do not have a predecessor school will not have this data. Since the idea of this project is to map likely outcomes for schools based on current and previous inspection data for the most part then I will filter these schools without prior outcomes from the dataset

```{r Filter out those without previous inspections}

(ofsted_filtered <- filter(ofsted, !is.na(Previous.full.inspection.overall.effectiveness)))

```

Next I'm going to make some exploratory charts. Hopefully confirming things I already suspect.

I'll start with the obvious one to do, prev inspection outcome vs current inspection outcome. I expect this to show an overall increase since Ofsteds published stats tell us this is the case.

```{r Chart previous outcome vs current outcome}

# This isn't going to work!

ggplot(ofsted_filtered) +
  geom_point(mapping = aes(x = Previous.full.inspection.overall.effectiveness, y = Overall.effectiveness))

```

That was a nice set of 16 points... On the bright side we know all possible outcome matches occur we just don't know which is happening a lot. So lets add some small variation to the points.

```{r Chars v2!}

# If we're lucky we'll get an idea of the trend now

ggplot(ofsted_filtered) +
  geom_point(mapping = aes(x = Previous.full.inspection.overall.effectiveness, y = Overall.effectiveness), position = "Jitter")
  
```
Much better! Now it is quite clear that most schools that were graded 2 previous stay at a two or move up to a 1. Most schools graded 3 moved up to a 2 and even the majority of 4 moved to a 2. There is some backsliding for schools that were 1 slipping back to a 2 but that is still a good outcome. However I would not present this chart to anyone. To make this a bit more user friendly I'm going to have to summarise the data and then produce a stacked bar chart from the outcomes I will produce two of these one will be a straight count the other will be proportion based.

This is a nice way of visualising tables:

```{r}
mosaicplot(xtabs(~ Overall.effectiveness + Previous.full.inspection.overall.effectiveness, data = ofsted_filtered))
```

```{r Summarise}

Ofsted_outcomes_summed <- ofsted_filtered %>% 
  group_by(Previous.full.inspection.overall.effectiveness, Overall.effectiveness) %>%
    summarise(n())

```
Now viewing this data we can see that the majority of the data points in the data set are 2-2 and 3-2, and 3-2 is in fact the larger number!

Next up is the first chart, previous outcome inspection vs current inspection as a stacked bar.

```{r Chart numero uno!}

ggplot(ofsted_filtered) +
  geom_bar(mapping = aes(x = Previous.full.inspection.overall.effectiveness, fill = as.character(Overall.effectiveness)))

```
It turns out I didn't need to summarise at all but it was good learning anyway. Now this chart clearly shows the majority of Previous inspections were 2 or 3 and of these the vast majority have ended up with a 2 or better outcome. But there are still some that need to be targeted for intervention to raise their outcomes as even 1's can become 4's. Hopefully the next chart will make it clearer.

```{r Chart numero dos!}

# Essentially the same code as the previous chart but just a small addition to change to a proportion view

ggplot(ofsted_filtered) +
  geom_bar(mapping = aes(x = Previous.full.inspection.overall.effectiveness, fill = as.character(Overall.effectiveness)), position = "Fill")

```
This project aims to identify those schools that are currently at any level of overall effectiveness, figure out what they are likely to achieve at their next inspection and then put in intervention to try and squeeze down the size of the 3 and 4 bands by turning them into 1's and 2's.

It might be a good idea to contrast this chart with that of Warwickshire

```{r Warwickshire chart}

warks_filtered <- ofsted_filtered %>%
  filter(Local.authority == "Warwickshire")

ofsted_filtered_chart <- mutate(ofsted_filtered, plotGroup = "National")

warks_filtered <- mutate(warks_filtered, plotGroup = "Warwickshire")

ofsted_filtered_chart <- ofsted_filtered_chart %>%
  add_row(warks_filtered)

ggplot(ofsted_filtered_chart) +
  geom_bar(mapping = aes(x = Previous.full.inspection.overall.effectiveness, 
                         fill = as.character(Overall.effectiveness)
                         ), 
           position = "Fill"
           ) +
  facet_wrap(~ plotGroup)

```
As you can see we have a larger proportion of 4's going to 3's than occurs in the national dataset. We also have no 1's going down to 4's.




Time to start working on a logistic regression of my dataset. Next we will do a test of logistic regression.

```{r First prep data ready for the regression test}

# The IDACI part had () as part of the name and that was causing a lot of issues so the following piece of code goes through all the column names of the dataframe and removes punctuation and spaces. Hopefully this is something that can be done earlier on, possibly when the files are read in at some point.

colnames(ofsted_filtered) <- str_replace_all(colnames(ofsted_filtered),"[:punct:]|[:space:]","")

ofsted_filtered <- mutate(ofsted_filtered, prev_out = as.integer(ifelse(Previousfullinspectionoveralleffectiveness == "NULL", -1, Previousfullinspectionoveralleffectiveness)))

ofsted_filtered <- mutate(ofsted_filtered, current_out = as.integer(ifelse(Overalleffectiveness == "NULL", -1, Overalleffectiveness)))

ofsted_filtered <- mutate(ofsted_filtered, IDACI = as.integer(ifelse(TheincomedeprivationaffectingchildrenindexIDACIquintile == "NULL", -1, TheincomedeprivationaffectingchildrenindexIDACIquintile)))

ofsted_filtered <- mutate(ofsted_filtered, num_pupils = as.integer(ifelse(Totalnumberofpupils == "NULL", -1, Totalnumberofpupils)))

ofsted_filtered_reg <- filter(ofsted_filtered, prev_out != -1 & current_out != -1 & IDACI != -1 & num_pupils != -1)

ofsted_filtered_reg <- mutate(ofsted_filtered_reg, pos_out = ifelse(current_out - prev_out <= 0,1,0))

ofsted_filtered_reg <- mutate(ofsted_filtered_reg, outcome_change = ifelse(Previousfullinspectionoveralleffectiveness == -1 | Overalleffectiveness == -1, -1, Overalleffectiveness - Previousfullinspectionoveralleffectiveness))

```

```{r Addtional filtering down to just those schools that have decreased by one step}

ofsted_filtered_reg_tree <- filter(ofsted_filtered_reg, outcome_change == 1)

ofsted_filtered_reg_tree$Ofstedphase <- as.factor(ofsted_filtered_reg_tree$Ofstedphase)

ofsted_filtered_reg_tree <- ofsted_filtered_reg_tree %>% 
  mutate(prev_out = factor(prev_out),
         curr_out = factor(current_out),
         IDACI = factor(IDACI),
         Academy = factor(Academy))

```

```{r Now test the regression}

# ofsted_filtered_reg$prev_out <- as.factor(ofsted_filtered_reg$prev_out)
ofsted_filtered_reg$Ofstedphase <- as.factor(ofsted_filtered_reg$Ofstedphase)

ofsted_filtered_reg <- ofsted_filtered_reg %>% 
  mutate(prev_out = factor(prev_out),
         curr_out = factor(current_out),
         IDACI = factor(IDACI),
         Academy = factor(Academy))

logit <- glm(pos_out ~ prev_out + IDACI + Academy + Ofstedphase + daysbetween, data = ofsted_filtered_reg, family = "binomial")

summary(logit)

exp(coef(logit))

# Tree based regression?

```
So on the basis of these results the most important precursor to a positive outcome is the previous inspection outcome and bizarrely if the school is a special school.

Another model would be tree based regression, so I will give that a go and see if I get similar results.

```{r Tree based regression test}

tree_test <- data_frame(ofsted_filtered_reg_tree)

tree <- rpart(current_out ~ prev_out, data = tree_test, method = "class")

printcp(tree)
plotcp(tree)
rsq.rpart(tree)
print(tree)
summary(tree)
plot(tree)
text(tree)

```

```{r Tree test 2!}

ofsted_tree_root <- ofsted_filtered %>%
  select(LAESTAB, Ofstedphase, Localauthority, Easting, Northing, Academy, DistrictAdministrativecode, LSOAcode, LACode, prev_out, current_out, IDACI, num_pupils)

ofsted_tree_root <- ofsted_tree_root %>% 
  mutate(prev_out = factor(prev_out),
         curr_out = factor(current_out),
         Ofstedphase = factor(Ofstedphase),
         IDACI = factor(IDACI),
         Academy = factor(Academy))

ofsted_tree_root <- ofsted_tree_root %>%
  filter(!(is.na(IDACI)) & !(is.na(num_pupils)))

ofsted_tree_sprout <- ofsted_tree_root %>%
  select(prev_out, curr_out, Ofstedphase, Academy, IDACI, num_pupils)

set.seed(061183)

ofsted_split <- initial_split(ofsted_tree_sprout, strata = prev_out)

ofsted_train <- training(ofsted_split)

ofsted_test <- testing(ofsted_split)



set.seed(310188)

ofsted_boot <- bootstraps(ofsted_train)
ofsted_boot

log_reg_mod <- logistic_reg() %>%
  set_engine("glm")

rf_mod <- rand_forest() %>%
  set_mode("classification") %>%
  set_engine("ranger")



ofsted_wf <- workflow() %>%
  add_formula(curr_out ~ .)

ofsted_wf



log_res_results <- ofsted_wf %>%
  add_model(log_reg_mod) %>%
  fit_resamples(
    resamples = ofsted_boot,
    control = control_resamples(save_pred = TRUE, verbose = TRUE)
  )

rf_results <- ofsted_wf %>%
  add_model(rf_mod) %>%
  fit_resamples(
    resamples = ofsted_boot,
    control = control_resamples(save_pred = TRUE, verbose = TRUE)
  )



out <- collect_metrics(rf_results)

view(out)

out2 <- rf_results %>%
  conf_mat_resampled()

view(out2)

rf_results %>%
  collect_predictions() %>%
  group_by(id) %>%
  roc_curve(curr_out, c(.pred_1, .pred_2, .pred_3, .pred_4)) %>%
  ggplot(aes(1 - specificity, sensitivity, color = id)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(show.legend = FALSE, alpha = 0.6, size = 1.2) +
  coord_equal()



ofsted_tree_leaves <- ofsted_wf %>%
  add_model(rf_mod) %>%
  last_fit(ofsted_split)

ofsted_tree_leaves



```
That was interesting but I want to do it again with a different variable

```{r Tree test 3!}

ofsted_tree_root <- ofsted_filtered %>%
  select(LAESTAB, Ofstedphase, Localauthority, Easting, Northing, Academy, DistrictAdministrativecode, LSOAcode, LACode, prev_out, current_out, IDACI, num_pupils)

ofsted_tree_root <- ofsted_tree_root %>% 
  mutate(prev_out = factor(prev_out),
         curr_out = factor(current_out),
         Ofstedphase = factor(Ofstedphase),
         IDACI = factor(IDACI),
         Academy = factor(Academy))

ofsted_tree_root <- ofsted_tree_root %>%
  filter(!(is.na(IDACI)) & !(is.na(num_pupils)))

ofsted_tree_root <- ofsted_tree_root %>%
  mutate(prev_good = ifelse(prev_out == "1" | prev_out == "2", "Y", "N"), 
         curr_good = ifelse(curr_out == "1" | curr_out == "2", "Y", "N"))

ofsted_tree_root <- ofsted_tree_root %>%
  mutate(prev_good = factor(prev_good), curr_good = factor(curr_good))


ofsted_tree_sprout <- ofsted_tree_root %>%
  select(prev_good, curr_good, Ofstedphase, Academy, IDACI, num_pupils)

set.seed(061183)

ofsted_split <- initial_split(ofsted_tree_sprout, strata = prev_good)

ofsted_train <- training(ofsted_split)

ofsted_test <- testing(ofsted_split)



set.seed(310188)

ofsted_boot <- bootstraps(ofsted_train)
ofsted_boot

log_reg_mod <- logistic_reg() %>%
  set_engine("glm")

rf_mod <- rand_forest() %>%
  set_mode("classification") %>%
  set_engine("ranger")



ofsted_wf <- workflow() %>%
  add_formula(curr_good ~ .)

ofsted_wf



log_res_results <- ofsted_wf %>%
  add_model(log_reg_mod) %>%
  fit_resamples(
    resamples = ofsted_boot,
    control = control_resamples(save_pred = TRUE, verbose = TRUE)
  )

rf_results <- ofsted_wf %>%
  add_model(rf_mod) %>%
  fit_resamples(
    resamples = ofsted_boot,
    control = control_resamples(save_pred = TRUE, verbose = TRUE)
  )

outlog <- collect_metrics(log_res_results)

out <- collect_metrics(rf_results)

view(out)

out2 <- rf_results %>%
  conf_mat_resampled()

view(out2)

rf_results %>%
  collect_predictions() %>%
  group_by(id) %>%
  roc_curve(curr_good, .pred_N) %>%
  ggplot(aes(1 - specificity, sensitivity, color = id)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(show.legend = FALSE, alpha = 0.6, size = 1.2) +
  coord_equal()



ofsted_tree_leaves <- ofsted_wf %>%
  add_model(rf_mod) %>%
  last_fit(ofsted_split)

ofsted_tree_leaves

collect_metrics(ofsted_tree_leaves)



```


```{r The fun zone! For tests and giggles!}

# Recreate Bostock Sankey diagram: http://bost.ocks.org/mike/sankey/
# Load energy projection data
URL <- paste0('https://cdn.rawgit.com/christophergandrud/networkD3/master/JSONdata/energy.json')
energy <- jsonlite::fromJSON(URL)

# Plot
sankeyNetwork(Links = energy$links, Nodes = energy$nodes, Source = 'source',
             Target = 'target', Value = 'value', NodeID = 'name',
             units = 'TWh', fontSize = 12, nodeWidth = 30)

# Colour links
energy$links$energy_type <- sub(' .*', '',
                               energy$nodes[energy$links$source + 1, 'name'])

sankeyNetwork(Links = energy$links, Nodes = energy$nodes, Source = 'source',
             Target = 'target', Value = 'value', NodeID = 'name',
             LinkGroup = 'energy_type', NodeGroup = NULL)



```

Testing out sankey plot but looks like it requires Shiny stuff that I have not learnt yet. So will continue on for now with other stuff!

```{r Sankey plot}
install.packages("sankey")
install.packages("networkD3")

library(sankey)
library(networkD3)

test_table <- tribble(
  ~ID, ~PrevOut, ~CurrOut, ~Value,
  #--/-/-/----
  1,1,1,100,
  2,1,2,200,
  3,1,3,300,
  4,1,4,400,
  5,2,1,200,
  6,2,2,400,
  7,2,3,600,
  8,2,4,800,
  9,3,1,300,
  10,3,2,600,
  11,3,3,900,
  12,3,4,1200,
  13,4,1,400,
  14,4,2,800,
  15,4,3,1200,
  16,4,4,1600)

test_table2 <- tibble(
  Col1 = "ID",
  Col2 = "PrevOut",
  Col3 = "CurrOut",
  Col4 = "Value"
)
testing1 <- sankeyNetwork(Links = test_table, Nodes = test_table2, Source = test_table$PrevOut, Target = test_table$CurrOut, units = test_table$Value)

sankeyNetworkOutput(testing1)

small_filtered <- ofsted_filtered %>%
  select(Previousfullinspectionoveralleffectiveness, Overalleffectiveness) %>%
  group_by(Previousfullinspectionoveralleffectiveness, Overalleffectiveness) %>%
  summarise(total = n())

new_small_filtered <- small_filtered %>%
  mutate(link = paste("'Source':",Previousfullinspectionoveralleffectiveness,",'Target':",Overalleffectiveness,",'Value':",total))

table_headers <- read.table(stringsAsFactors = FALSE, textConnection(
  "Previousfullinspectionoveralleffectiveness Overalleffectiveness")
)

testing1 <- sankeyNetwork(Links = new_small_filtered$link, Nodes = table_headers, Source = c(1:16), Target = c(1:16), units = new_small_filtered$total)

ggplot(small_filtered) +
  geom_ribbon(mapping = )
  
sankey_data <- make_sankey(edges = small_filtered)

sankey(sankey_data)

```